Cosmic web alignments with the shape, angular momentum and peculiar velocities of dark matter halos

### 2014-02-13 paper/main.tex
comments from Sergio

\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}} 
\newcommand{\muavg}{\vert\langle\cos\theta\rangle\vert}
% definition to produce a "less than or similar to" symbol
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}

@@

a \cite{Trowland2013} presented a parameterization for the $\mu$
distribution in the case of angular momentm alignment based on
theoretical considerations by \cite{Lee2005} (Eq. \ref{eq:distro} in
the Appendix A). Under this parameterization it was found a unique
correspondece between the full shape of the $\vert\mu\vert$
distribution and its average. In this paper we use that result to only
present the results for the average $\langle\vert\mu\vert\rangle$.

@@

\begin{enumerate}
\item Halos with $(\hat{s}\cdot\hat{e}_1)^2> (\hat{s}\cdot\hat{e}_2)^2$
  and $(\hat{s}\cdot\hat{e}_1)^2> (\hat{s}\cdot\hat{e}_2)$, which can
  be considered to aligned mostly along $\hat{e}_1$.
\item Halos with $(\hat{s}\cdot\hat{e}_2)^2> (\hat{s}\cdot\hat{e}_2)^2$
  and $(\hat{s}\cdot\hat{e}_2)^2> (\hat{s}\cdot\hat{e}_3)^2$, which can
  be considered to aligned mostly along $\hat{e}_2$.
\item Halos with $(\hat{s}\cdot\hat{e}_3)^2> (\hat{s}\cdot\hat{e}_1)^2$
  and $(\hat{s}\cdot\hat{e}_3)^2> (\hat{s}\cdot\hat{e}_1)^2$, which can
  be considered to aligned mostly along $\hat{e}_2$.
\end{enumerate}

@@

The shape alignment and the Vweb (upper row, left column) gives a
different perspective. Firstly, there seems to be little evidence
for an alignment for masses below $10^{11}-10^{12}$\hMsun, depending
on the grid resolution. Secondly, the alignment at higher masses goes
along the first eigenvector $\vec{e}_{1}$ meaning that they mostly lie
perpendicular to the filaments and sheets. In the discussion section
we clarify this result that at first might seem puzzling.

For fhe angular momentum alignment and the Tweb (middle row, right
column) we find that at low masses $<10^{12}$\hMsun no evidence for
any alignment. At higher masses $>10^{12}$\hMsun there is weak signal
of preferential alignment along the first and second eigenvectors,
between $35\%$ to $45\%$ of the halos are aligned with respect to
$\vec{e}_1$ and $\vec{e}_2$. Correspondingly, between $10\%$ to $20\%$
of the halos are aligned along respect to $\vec{e}_{3}$. This means
that most of the halos are perpendicular to the filaments and do not
have a clear alignment with respect to walls.

The angular momentum in the Vweb (middle row, left column) presents a
a signal of alignment along the second eigenvector $\hat{e}_{2}$,
between $45\%$ to $60\%$ of the halos are aligned along that
direction. While there is a minority of halos aligned with
$\vec{e}_{1}$.  There a clear change in trends around
$10^{11}$\hMsun-$10^{12}$\hMsun depending on the grid resolution;
below that mass range there is no evidence for alignment while at
higher masses appear all the trends we describe. This means that most
of the halos tend to lie along walls but do not have a clear alignment
with respect to filaments.

The peculiar velocities show a weak but consistent alignment along the
third eigenvector $\vec{e}_{3}$ of the Tweb for all masses below
$10^{13.0}$\hMsun-$10^{13.5}$\hMsun depending on the grid
resolution. $45\%$ of the halos are aligned while only $25\%$ are
aligned along the first eigenvector $\vec{e}_1$. This suggests that
halos tend to move along filaments and parallel to the walls, except
at higher masses where the alignments get randomized.  In contrast,
the peculiar velocities with respect to the Vweb show the same trend
but weaker and only for low mass halos $<10^{12}$.

@@

We do not find a strong signal for the alignment of spin and the
cosmic web. A significant correlation signal has been recently
measured between the spin and the cosmic vorticity
\citep{Libeskind2013a,Laigle2013}. This two facts are consistent with
spin being produced by the non-symmetric part of the shear and tidal
tensors.

>>>>

\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$~}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}} 
\newcommand{\muavg}{\vert\langle\cos\theta\rangle\vert}
% definition to produce a "less than or similar to" symbol
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}

@@

\cite{Trowland2013} presented a parameterization for the $\mu$
distribution in the case of angular momentm alignment based on
theoretical considerations by \cite{Lee2005} (Eq. \ref{eq:distro} in
the Appendix A). Under this parameterization it was found a unique
correspondece between the full shape of the $\vert\mu\vert$
distribution and its average. In this paper we use that result to only
present the results for the average $\langle\vert\mu\vert\rangle$.

@@

\begin{enumerate}
\item Halos with $(\hat{s}\cdot\hat{e}_1)^2> (\hat{s}\cdot\hat{e}_2)^2$
  and $(\hat{s}\cdot\hat{e}_1)^2> (\hat{s}\cdot\hat{e}_3)$, which can
  be considered to aligned mostly along $\hat{e}_1$.
\item Halos with $(\hat{s}\cdot\hat{e}_2)^2> (\hat{s}\cdot\hat{e}_1)^2$
  and $(\hat{s}\cdot\hat{e}_2)^2> (\hat{s}\cdot\hat{e}_3)^2$, which can
  be considered to aligned mostly along $\hat{e}_2$.
\item Halos with $(\hat{s}\cdot\hat{e}_3)^2> (\hat{s}\cdot\hat{e}_1)^2$
  and $(\hat{s}\cdot\hat{e}_3)^2> (\hat{s}\cdot\hat{e}_2)^2$, which can
  be considered to aligned mostly along $\hat{e}_3$.
\end{enumerate}

@@

The shape alignment and the Vweb (upper row, left column) gives a
different perspective. Firstly, there seems to be little evidence
for an alignment for masses below $10^{11}-10^{12}\hMsun$, depending
on the grid resolution. Secondly, the alignment at higher masses goes
along the first eigenvector $\vec{e}_{1}$ meaning that they mostly lie
perpendicular to the filaments and sheets. In the discussion section
we clarify this result that at first might seem puzzling.

For the angular momentum alignment and the Tweb (middle row, right
column) we find that at low masses $<10^{12}\hMsun$ no evidence for
any alignment. At higher masses $>10^{12}\hMsun$ there is weak signal
of preferential alignment along the first and second eigenvectors,
between $35\%$ to $45\%$ of the halos are aligned with respect to
$\vec{e}_1$ and $\vec{e}_2$. Correspondingly, between $10\%$ to $20\%$
of the halos are aligned along respect to $\vec{e}_{3}$. This means
that most of the halos are perpendicular to the filaments and do not
have a clear alignment with respect to walls.

The angular momentum in the Vweb (middle row, left column) presents a
a signal of alignment along the second eigenvector $\hat{e}_{2}$,
between $45\%$ to $60\%$ of the halos are aligned along that
direction. While there is a minority of halos aligned with
$\vec{e}_{1}$.  There a clear change in trends around
$10^{11}\hMsun-10^{12}\hMsun$ depending on the grid resolution;
below that mass range there is no evidence for alignment while at
higher masses appear all the trends we describe. This means that most
of the halos tend to lie along walls but do not have a clear alignment
with respect to filaments.

The peculiar velocities show a weak but consistent alignment along the
third eigenvector $\vec{e}_{3}$ of the Tweb for all masses below
$10^{13.0}\hMsun-10^{13.5}\hMsun$ depending on the grid
resolution. $45\%$ of the halos are aligned while only $25\%$ are
aligned along the first eigenvector $\vec{e}_1$. This suggests that
halos tend to move along filaments and parallel to the walls, except
at higher masses where the alignments get randomized.  In contrast,
the peculiar velocities with respect to the Vweb show the same trend
but weaker and only for low mass halos $<10^{12}$.

@@

We do not find a strong signal for the alignment of spin and the
cosmic web. A significant correlation signal has been recently
measured between the spin and the cosmic vorticity
\citep{Libeskind2013b,Laigle2013}. This two facts are consistent with
spin being produced by the non-symmetric part of the shear and tidal
tensors.
